### 2021-04-13 pldi21/abstract.tex
updated to latest style files, added required location data

\begin{abstract}
    Text editing is powerful, but 
    some types of expressions are more naturally represented and manipulated
    graphically.
    Examples include expressions that compute 
    colors,
    music,
    animations,
    tabular data, 
    plots,
    GUI widgets, 
    mathematical diagrams, 
    and other domain-specific data structures.
    This paper introduces \emph{live literals}, or \emph{livelits}, 
    which allow
    clients to fill holes of types like these   
    by directly manipulating a     
    user-defined GUI embedded persistently into code. 
    Uniquely, livelits are \emph{compositional}: a livelit GUI can itself embed 
    spliced expressions, which are typed, lexically scoped, enjoy full editor support and can in turn embed 
    other livelits. 
    Livelits are also uniquely \emph{live}: 
    a livelit can provide {continuous feedback} about the run-time implications 
    of the client's choices 
    even when splices mention bound variables, 
    because the system continuously gathers closures associated with the 
    hole that the livelit is tasked with filling.
    We integrate livelits into Hazel, 
    a live programming environment able to 
    typecheck and run programs with holes, 
    and describe several case studies that exercise these novel capabilities. 
    We then define a minimal typed livelit calculus, which precisely 
    specifies how livelits operate as 
    live graphical macros.     
    The metatheory of macro expansion has been mechanized in Agda.
\end{abstract}

>>>>

\begin{abstract}
    Text editing is powerful, but 
    some types of expressions are more naturally represented and manipulated
    graphically.
    Examples include expressions that compute 
    colors,
    music,
    animations,
    tabular data, 
    plots,
    diagrams, 
    and other domain-specific data structures.
    This paper introduces \emph{live literals}, or \emph{livelits}, 
    which allow
    clients to fill holes of types like these   
    by directly manipulating a     
    user-defined GUI embedded persistently into code. 
    Uniquely, livelits are \emph{compositional}: a livelit GUI can itself embed 
    spliced expressions, which are typed, lexically scoped, and can in turn embed 
    other livelits. 
    Livelits are also uniquely \emph{live}: 
    a livelit can provide {continuous feedback} about the run-time implications 
    of the client's choices 
    even when splices mention bound variables, 
    because the system continuously gathers closures associated with the 
    hole that the livelit is filling.
    We integrate livelits into Hazel, 
    a live hole-driven programming environment, 
    and describe case studies that exercise these novel capabilities. 
    We then define a simply typed livelit calculus, which  
    specifies how livelits operate as 
    live graphical macros.     
    The metatheory of macro expansion has been mechanized in Agda.
\end{abstract}

### 2021-04-13 pldi21/discussion.tex
updated to latest style files, added required location data

The authors would like to thank
Pavel Panchekha and 
the anonymous POPL 2021 and PLDI 2021 referees 
for thoughtful feedback that has significantly improved this paper. 
This material is based upon work supported by the National Science Foundation 
under Grant No. 1814900 and 1817145.

>>>>

The authors thank
Pavel Panchekha and 
 anonymous referees at TyDe 2019, POPL 2021 and PLDI 2021  
for thoughtful feedback that has significantly improved this paper. 
This material is based upon work supported by the National Science Foundation 
under Grant No. 1814900 and 1817145.

### 2021-04-13 pldi21/intro.tex
update grade-cutoffs figure + tweak subfigure sizes to make same height
fixed flow around fig 1
updated to latest style files, added required location data

Consider, as a simple example, a record type
classifying RGBA-encoded colors. 
It is possible to select a particular color by entering
an expression of this type in a text editor, e.g. \li{\{ r: 57, g: 107, b: 57, a: 92 \}}. 
The problem with this \emph{de facto} textual user interface for color selection is that 
it offers no live feedback about which color has been expressed 
and limited editing affordances for choosing a different color.
Analagous critiques apply to strictly textual user interfaces for 
countless other data structures,
e.g. vector graphics,
animation parameters,
musical sequences,
audio filters,
board game states, 
GUI widgets and layouts,
tabular numeric data, 
plots,
geospatial data, 
neural network diagrams, 
biological neuron models, 
and mathematical diagrams.

Practitioners in domains where manipulating data of types like these is 
a central activity 
have largely eschewed general-purpose programming environments 
in favor of specialized graphical end-user applications, e.g. %
image and video editors, music composition software, level design tools, 
and bespoke GUIs written by students or lab technicians.
This is in large part because these applications 
take seriously the need for live feedback, domain-specific 
non-textual data representations, and 
direct manipulation affordances, 
e.g. color palettes, visual timelines, plots, and maps.

@@

\begin{figure*}
  \begin{minipage}[t]{0.37\textwidth}
    \begin{subfigure}[t]{\linewidth}
    \begin{snugshade}
      \vspace*{-2mm}
      \caption{\textbf{Prior Work:} Graphite \cite{Graphite}}
       \label{fig:graphite}
      \vspace*{1mm}
     \end{snugshade}
      \vspace*{-1mm}
      \includegraphics[width=\linewidth]{graphite-color-palette-green.png}
      \vspace*{-5mm}
    \end{subfigure}
    \hspace{8mm}
    \begin{subfigure}[t]{\linewidth}
     \begin{snugshade}
      \vspace*{-2mm}
      \caption{\textbf{This Paper:} Livelits are live and compositional}
    \label{fig:color}
      \vspace*{1mm}
     \end{snugshade}
      \vspace*{-1mm}
      \includegraphics[width=\linewidth]{slider-color-livelits.png}
    \end{subfigure}
  \end{minipage}
  \hspace{10.5mm}
  \begin{subfigure}[t]{0.50\textwidth}
  \begin{snugshade}
   \vspace*{-2mm}
    \caption{\textbf{Case Study}: Grading with Livelits}
    \label{fig:grading}
    \vspace*{1mm}
     \end{snugshade}
    \vspace*{-1mm}
    \includegraphics[width=\linewidth]{grade-cutoff-livelit.png}
  \end{subfigure}
  \vspace{-4mm}
   \caption{Introductory Examples}
%   \vspace*{-2mm}
\end{figure*}

>>>>

Consider, as a simple example, a record type
classifying RGBA-encoded colors.
It is possible to select a particular color by entering
an expression of this type in a text editor, e.g. \li{\{ r: 57, g: 107, b: 57, a: 92 \}}.
The problem with this \emph{de facto} textual user interface for color selection is that
it offers no live feedback about which color has been expressed
and limited editing affordances for choosing a different color.
Analagous critiques apply to strictly textual user interfaces for
countless other data structures,
e.g. audio filters, 
vector graphics,
% animation parameters,
% musical sequences,
board game states,
% GUI widgets and layouts,
tabular numeric data,
% plots,
geospatial data,
% neural network diagrams,
neurobiological circuits,
and mathematical diagrams.

Practitioners in domains where manipulating data of types like these is
a central activity
have largely eschewed general-purpose programming environments
in favor of specialized graphical end-user applications, e.g. %
image and video editors, music composition software, level design tools,
and bespoke GUIs written by lab technicians.
This is in large part because these applications
take seriously the need for live feedback, domain-specific
non-textual data representations, and
direct manipulation affordances,
e.g. checkboxes, sliders, color palettes, visual timelines, interactive plots, and maps.

@@

\begin{figure*}
  \begin{minipage}[t]{0.38\textwidth}
    \begin{subfigure}[t]{\linewidth}
    \begin{snugshade}
      \vspace*{-2mm}
      \caption{\textbf{Prior Work:} Graphite \cite{Graphite}}
       \label{fig:graphite}
      \vspace*{1mm}
     \end{snugshade}
      \vspace*{-1mm}
      \includegraphics[width=\linewidth]{graphite-color-palette-green.png}
      \vspace*{-5mm}
    \end{subfigure}
    \hspace{8mm}
    \begin{subfigure}[t]{\linewidth}
     \begin{snugshade}
      \vspace*{-2mm}
      \caption{\textbf{This Paper:} Livelits are live and compositional}
    \label{fig:color}
      \vspace*{1mm}
     \end{snugshade}
      \vspace*{-1mm}
      \includegraphics[width=\linewidth]{slider-color-livelits.png}
    \end{subfigure}
  \end{minipage}
  \hspace{10.5mm}
  \begin{subfigure}[t]{0.55\textwidth}
  \begin{snugshade}
   \vspace*{-2mm}
    \caption{\textbf{Case Study}: Grading with Livelits}
    \label{fig:grading}
    \vspace*{1mm}
     \end{snugshade}
    \vspace*{-1mm}
    \includegraphics[width=\linewidth]{grade-cutoff-livelit.png}
  \end{subfigure}
  \vspace{-5mm}
   \caption{Introductory Examples}
%   \vspace*{-2mm}
\end{figure*}
